---
title: "Fire Sensitivity"
author: "Alec Dyer"
date: "October 20, 2024"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Libraries & Data

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
load('fire_data_filtered_qmd_v3.Rdata')
```

# FOFEM

>=0.75 FOFEM_PMORT as sensitive, <0.75 to >0.25 as tolerant, and < 0.25 as resistant

```{r}
fofem_qmd <- read_csv('FOFEM_DIACLASS.csv') %>% select(c('SPECIES_ABV', 'DIACLASS', 'FOFEM_PMORT_6FT_SFL'))
```

```{r}
df.fires.traits <- df.fires.traits %>%
  mutate(qmd = qmd / 10) %>% # scale
  mutate(qmd_in = qmd / 2.54) # to inches
```

```{r}
# Match FOFEM values based on species and diameter (inches)
df.fires.traits.fofem <- df.fires.traits %>%
  mutate(DIACLASS = ifelse(qmd_in >= 20, '>20', '')) %>%
  mutate(DIACLASS = ifelse(qmd_in < 20, '10TO20', DIACLASS)) %>%
  mutate(DIACLASS = ifelse(qmd_in < 10, '1TO10', DIACLASS)) %>%
  merge(fofem_qmd, by.x=c("fortypba", "DIACLASS"), by.y=c("SPECIES_ABV", "DIACLASS")) %>%
  mutate(sensitivity_fofem = cut(FOFEM_PMORT_6FT_SFL, breaks=c(0, 0.25, 0.75, Inf), labels=c('Resistant', 'Tolerant', 'Sensitive'))) %>%
  mutate(sensitivity_fofem = factor(sensitivity_fofem, levels=c('Sensitive', 'Tolerant', 'Resistant'))) %>%
  mutate(mortality = factor(mortality, levels=c('First-order','Delayed','Survived')))
```


```{r fig.width=5, fig.height=2}
df.fires.traits.fofem %>%
  filter(rdnbr > 0) %>%
  mutate(age_class = cut(age, breaks=c(-1, 80, 160, Inf), labels=c('Young', 'Adult', 'Mature'))) %>%
  group_by(sensitivity_fofem, age_class, mortality) %>%
  summarise(count=n()) %>%
  mutate(perc = count/sum(count)*100) %>%
  ggplot(aes(x=perc, y=age_class, fill=mortality)) +
    geom_bar(stat='identity', position='dodge', width=0.5, color='black') +
    scale_fill_manual(values=c('#E1E1E1', '#ff7f00', '#267300')) +
    ylab('Sensitivity Class') + 
    xlab('Percentage (%)') +
    theme_cowplot() +
    theme(
        axis.text.x=element_text(size=10),
        axis.text.y = element_text( size = 10 ),
        axis.title = element_text( size = 12),
        strip.text = element_text(size = 10),
        legend.position="bottom",
        legend.justification = "center"
      ) + 
    guides(fill=guide_legend(title="")) +
    facet_wrap(~sensitivity_fofem)

ggsave2('figures/change_by_firetrait_ageclass_rdnbr0_fofem.png')
```








