---
title: "Forest Classification"
author: "Alec Dyer"
date: "January 24, 2024"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')
```

```{r message=F, include=F}
library(dplyr)
library(raster)
library(terra)
library(tidyterra)
library(sf)
library(rgdal)
library(stats)
library(ggplot2)
library(caret)
library(MASS)
library(tidymodels)
library(rpart.plot)
library(vip)
library(xgboost)
library(randomForest)
library(rfUtilities)
library(ROSE)
library(splitTools)
library(matrixStats)
library(here)
library(parallel)
library(corrplot)
library(car)
library(pROC)
```

# Inputs


```{r eval=F}
fires <- c('Riverside','BeachieCreek','Lionshead','HolidayFarm','ArchieCreek')

df.img <- data.frame(
  Fire = c('Riverside','Riverside','Riverside','Riverside','Riverside',
           'BeachieCreek','BeachieCreek','BeachieCreek','BeachieCreek','BeachieCreek',
           'Lionshead','Lionshead','Lionshead','Lionshead','Lionshead',
           'HolidayFarm','HolidayFarm','HolidayFarm','HolidayFarm','HolidayFarm',
           'ArchieCreek','ArchieCreek','ArchieCreek','ArchieCreek','ArchieCreek'),
  
  Year = c('2020_pre', 2020, 2021, 2022, 2023,
           '2020_pre', 2020, 2021, 2022, 2023,
           '2020_pre', 2020, 2021, 2022, 2023,
           '2020_pre', 2020, 2021, 2022, 2023,
           '2020_pre', 2020, 2021, 2022, 2023),
  
  Stack = c(here('imagery_normal_processed/riverside_2020_prefire.tif'),
            here('imagery_normal_processed/riverside_2020.tif'),
            here('imagery_normal_processed/riverside_2021.tif'),
            here('imagery_normal_processed/riverside_2022.tif'),
            here('imagery_normal_processed/riverside_2023.tif'),
            here('imagery_normal_processed/beachie_creek_2020_prefire.tif'),
            here('imagery_normal_processed/beachie_creek_2020.tif'),
            here('imagery_normal_processed/beachie_creek_2021.tif'),
            here('imagery_normal_processed/beachie_creek_2022.tif'),
            here('imagery_normal_processed/beachie_creek_2023.tif'),
            here('imagery_normal_processed/lionshead_2020_prefire.tif'),
            here('imagery_normal_processed/lionshead_2020.tif'),
            here('imagery_normal_processed/lionshead_2021.tif'),
            here('imagery_normal_processed/lionshead_2022.tif'),
            here('imagery_normal_processed/lionshead_2023.tif'),
            here('imagery_normal_processed/holiday_farm_2020_prefire.tif'),
            here('imagery_normal_processed/holiday_farm_2020.tif'),
            here('imagery_normal_processed/holiday_farm_2021.tif'),
            here('imagery_normal_processed/holiday_farm_2022.tif'),
            here('imagery_normal_processed/holiday_farm_2023.tif'),
            here('imagery_normal_processed/archie_creek_2020_prefire.tif'),
            here('imagery_normal_processed/archie_creek_2020.tif'),
            here('imagery_normal_processed/archie_creek_2021.tif'),
            here('imagery_normal_processed/archie_creek_2022.tif'),
            here('imagery_normal_processed/archie_creek_2023.tif'))
)

df.pts <- c(
  Riverside = here('samples/riverside_pts.shp'),
  BeachieCreek = here('samples/beachiecreek_pts.shp'),
  Lionshead = here('samples/lionshead_pts.shp'),
  HolidayFarm = here('samples/holidayfarm_pts.shp'),
  ArchieCreek = here('samples/archiecreek_pts.shp')
)

df.pif <- c(
  Riverside = here('samples/riverside_pts_pif.shp'),
  BeachieCreek = here('samples/beachiecreek_pts_pif.shp'),
  Lionshead = here('samples/lionshead_pts_pif.shp'),
  HolidayFarm = here('samples/holidayfarm_pts_pif.shp'),
  ArchieCreek = here('samples/archiecreek_pts_pif.shp')
)
```

# Data Prep

```{r}
pts.0 <- apply(df.img %>% filter(Year=='2020_pre'), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pif[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))

pts.1 <- apply(df.img %>% filter(Year==2020), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pif[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))

pts.2 <- apply(df.img %>% filter(Year==2021), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pif[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))

pts.3 <- apply(df.img %>% filter(Year==2022), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pts[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))

pts.4 <- apply(df.img %>% filter(Year==2023), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pif[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))

pts.3.pif <- apply(df.img %>% filter(Year==2022), 1,  FUN=function(row) {
  # open data
  r <- rast(row['Stack'])
  pts <- vect(df.pif[row['Fire']]) %>% terra::project('epsg:32610') %>% 
    dplyr::select(US_L4CODE, struccond, loc, Uncertain, Forest, Shadow, ForType, PercCov, Cover)
  
  # extract points
  ext <- terra::extract(r, pts, raw=FALSE)
  # join dfs
  pts.join <- cbind(pts, ext) %>% as.data.frame()
  # add fire name
  pts.join$fire <- row['Fire']
  pts.join$year <- row['Year']
  
  return(pts.join)
  
}) %>% bind_rows() %>% mutate(Forest_B = as.factor(Forest))
```




# Supervised Classification

## Feature Analysis

```{r}
# combine point data into one
pts.comb <- rbind(pts.0, pts.1, pts.2, pts.3, pts.4) %>% drop_na() %>% dplyr::select(Forest_B, R, G, B, NIR, NDVI, GNDVI, BRGHT, SDB, RDNBR)
pts.comb$Forest_B <- as.numeric(pts.comb$Forest_B)
```

```{r eval=F}
ggplot(gather(pts.comb %>% dplyr::select(-Forest_B)), aes(value)) + 
    geom_histogram(bins = 20) + 
    facet_wrap(~key, scales = 'free_x')
```

```{r warning=F, eval=F}
# multicollinearity matrix
pts.comb.corr <- cor(pts.comb %>% dplyr::select(-Forest_B))
```

```{r eval=F}
(g.a <- corrplot.mixed(pts.comb.corr, lower='number', number.cex=0.75))
```

```{r eval=F}
# create linear model
model <- lm(Forest_B ~ ., data=pts.comb %>% dplyr::select(-c(R,G,B)))
# run VIF on model
vif(model)
```


## Modelling

### Data Preparation

```{r}
# drop rows with NA values
pts.0 <- pts.0 %>% drop_na(any_of('SDB'))
pts.1 <- pts.1 %>% drop_na(any_of('SDB'))
pts.2 <- pts.2 %>% drop_na(any_of('SDB'))
pts.3 <- pts.3 %>% drop_na(any_of('SDB'))
pts.3.pif <- pts.3.pif %>% drop_na(any_of('SDB'))
pts.4 <- pts.4 %>% drop_na(any_of('SDB'))
```

Get point data in 2020, 2021, and 2023 for non-burned (nb) point locations

```{r}
pts.nb <- bind_rows(pts.0, pts.1, pts.2, pts.3.pif, pts.4)

# combine NB observations with the original training data
pts.comb <- bind_rows(pts.0, pts.1, pts.2, pts.3, pts.4)
```

```{r}
# split non-burned points in 2:1 ratio
pts.nb.a <- pts.nb[1:588,] # 1:467
pts.nb.b <- pts.nb[589:882,] # 468:701

# create combined data using only half of the non-burned points
pts.comb.half <- bind_rows(pts.3, pts.nb.a)
```

### Modeling for 2022 Data

```{r}
s = 123 # setting seed
# features to train on based on VIF analysis
feats <- c('Forest_B', 'NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire')
# subset df
pts.stack.3.s <- pts.3 %>% dplyr::select(feats)
# 10 stratified folds
folds <- create_folds(pts.stack.3.s$Forest_B, k = 10, seed = s, m_rep = 1, shuffle=FALSE)
# save accuracy scores for each fold
cv_acc <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_acc) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')
# save AUC scores for each fold
cv_auc <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_auc) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')
# object to save variable importance results to
cv_imp <- data.frame(matrix(ncol = length(feats)-1, nrow = 10))
colnames(cv_imp) <- feats[-1]
# save internal rf model mean decrease in accuracy for each feature
cv_imp_rf <- data.frame(matrix(ncol = length(feats)-1, nrow = 10))
colnames(cv_imp_rf) <- feats[-1]

if (exists("cv.cm")) {
  remove(cv.cm)
}

# loop over each set of train and test folds
for (i in seq_along(folds)) {
  train <- pts.stack.3.s[folds[[i]], ]
  test <- pts.stack.3.s[-folds[[i]], ]
  
  # perform over-sampling
  train <- ovun.sample(Forest_B~., data=train, method='over')$data
  
  # train DT model
  dt.mdl <- decision_tree() %>%
    set_engine("rpart") %>%
    set_mode("classification") %>% 
    fit(Forest_B ~ ., data = train)
  
  # train RF model
  rf.mdl <- randomForest(Forest_B ~ .,
    data = train,
    importance = TRUE,
    ntree = 1000, # n rounds
  )
  
  # train BRT model
  brt.mdl <- boost_tree(
    mode = "classification",
    trees = 1000, #nrounds
  ) %>%
    set_engine("xgboost", verbose=0) %>%
    fit(Forest_B ~ ., data = train)
  
  # Make predictions on the testing data
  dt.p <- as.matrix(dt.mdl %>% predict(test))
  rf.p <- as.matrix(rf.mdl %>% predict(test))
  brt.p <- as.matrix(brt.mdl %>% predict(test))
  
  # confusion matrices
  dt.cm <- confusionMatrix(table(dt.p, test$Forest_B), positive='1')
  rf.cm <- confusionMatrix(table(rf.p, test$Forest_B), positive='1')
  brt.cm <- confusionMatrix(table(brt.p, test$Forest_B), positive='1')
  
  # add random forest confusion matrix
  if (exists("cv.cm")) {
    cv.cm <- cv.cm + rf.cm$table
  } else {
    cv.cm <- rf.cm$table
  }
  
  # save accuracy scores from confusion matrices
  cv_acc[i,'Decision Tree'] <- dt.cm$byClass[11]
  cv_acc[i,'Random Forest'] <- rf.cm$byClass[11]
  cv_acc[i,'Boosted Regression Tree'] <- brt.cm$byClass[11]
  
  # save AUC scores from confusion matrices
  test$Forest <- as.numeric(test$Forest_B)
  cv_auc[i,'Decision Tree'] <- auc(dt.p, test$Forest)
  cv_auc[i,'Random Forest'] <- auc(rf.p, test$Forest)
  cv_auc[i,'Boosted Regression Tree'] <- auc(brt.p, test$Forest)
  
  # save feature importance 
  var_imp <- vip::vip(rf.mdl)
  df_imp <- as.data.frame(var_imp$data)
  df_imp <- df_imp[match(feats[-1], df_imp$Variable),]
  cv_imp[i,] <- as.vector(df_imp$Importance)
  
  # save permuatation feature importance
  cv_imp_rf[i,] <- rf.mdl$importance[,3]
}
```

```{r fig.cap='Figure 6: Boxplot showing the distribution of the cross-validated balanced accuracy from the supervised forest classification models.'}
(g1 <- boxplot(cv_acc, horizontal=F, ylab='Balanced Accuracy', xlab='Model', main='Cross-Validated Balanced Accuracy'))
(g1.auc <- boxplot(cv_auc, horizontal=F, ylab='AUC', xlab='Model', main='Cross-Validated AUC Score'))
```

```{r}
(g2 <- cv_acc %>%
  gather('Model', 'Acc', `Decision Tree`:`Boosted Regression Tree`) %>%
  ggplot( aes(x=Model, y=Acc, fill=Model)) +
    geom_boxplot() +
    # scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    # theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    # ggtitle("A boxplot with jitter") +
    xlab("Model") + ylab('Balanced Accuracy'))
```


```{r}
feats.ordered <- colMedians(cv_imp %>% as.matrix()) %>% sort(decreasing = FALSE) %>% names()

(g3 <- cv_imp %>%
  gather('Feature', 'Value') %>%
  mutate(Value=as.numeric(Value)) %>%
  ggplot(aes(y=factor(Feature, level = feats.ordered), x=Value)) + geom_boxplot() + ylab('Feature') + xlab('Importance'))
```

```{r}
feats.ordered <- colMedians(cv_imp_rf %>% as.matrix()) %>% sort(decreasing = FALSE) %>% names()

(g4 <- cv_imp_rf %>%
  gather('Feature', 'Value') %>%
  mutate(Value=as.numeric(Value)) %>%
  ggplot(aes(y=factor(Feature, level = feats.ordered), x=Value)) + geom_boxplot() + ylab('Feature') + xlab('Decrease in Accuracy'))
```

### Sentitivity Test

To determine how many training samples are needed to develop accurate classification models for this region, a sensitivity test will be performed by which the number of training samples will vary from low (n=7) to high (n=147) with increments of 10 observations. A held out testing set (n=17) will be used to validate

```{r eval=F}
sq <- seq( 20, 700, 20) # seq(37, 147, 10)

# save error scores for each fold
cv_acc_sens <- data.frame(matrix(ncol = length(sq), nrow = 10))
colnames(cv_acc_sens) <- sq %>% as.character()

# loop over each set of train and test folds
for (i in seq_along(sq)) {
  df.t <- pts.stack.3.s %>% sample_n(sq[i])
  
  folds <- createFolds(df.t$Forest_B, list=TRUE, returnTrain = TRUE)
  
  # loop over each set of train and test folds
  for (j in seq_along(folds)) {
    df.t <- df.t[folds[[j]], ]
    part <- createDataPartition(df.t$Forest_B, p=0.7)
    train <- df.t[part$Resample1, ] %>% drop_na()
    test <- df.t[-part$Resample1, ] %>% drop_na()
    
    # train RF model
    rf.mdl <- randomForest(Forest_B ~ .,
      data = train,
      importance = TRUE,
      ntree = 1000, # n rounds
    )
    
    # Make predictions on the testing data
    rf.p <- as.matrix(rf.mdl %>% predict(test))
  
    # confusion matrices
    rf.cm <- confusionMatrix(table(factor(rf.p, levels=0:1), 
      factor(test$Forest_B, levels=0:1)))
    
    # save accuracy scores from confusion matrices
    cv_acc_sens[j, sq[i] %>% as.character()] <- rf.cm$byClass[11]
  }
  
}
```


```{r eval=F}
boxplot(cv_acc_sens, horizontal=F, ylab='Balanced Accuracy', xlab='# Samples', main='Cross-Validated Balanced Accuracy')
```

```{r eval=F}
library(OOBCurve)
```

```{r eval=F}
rf.mdl.3 <- randomForest(Forest_B ~ .,
  data = pts.stack.3.s %>% dplyr::select(-fire),
  importance = TRUE,
  ntree = 100, # n rounds
  keep.inbag=TRUE
)

sonar.task = makeClassifTask(data = pts.stack.3.s %>% dplyr::select(-fire), target = "Forest_B")

oob <- OOBCurve(rf.mdl.3, data=pts.stack.3.s, task=sonar.task)

plot(oob$auc)
```



### Model Performance on Temporal Imagery

NAIP imagery is only available in 2022. However, forest cover classifications need to be accurate across imagery over time (i.e., in 2020, 2021, and 2023). Therefore, model accuracy will also be tested on imagery from the other years. Since land cover can change over time, only observations with no disturbances (n=34) will be used for validation.

Balanced accuracy results on the temporal imagery are shown in Figure 8. Accuracy on temporal imagery can be very low when training on only the 2022 data (Without Combined Data group in Figure 8). However, if the model is also trained on the non-disturbed data from 2020, 2021, and 2023, model accuracy improves drastically for all models (With Combined Data Group in Figure 8). The Random Forest model is still the best performing model.


```{r}
s = 123 # setting seed

pts.comb.s <- pts.comb %>% dplyr::select(feats) 
pts.comb.half.s <- pts.comb.half %>% dplyr::select(feats)

# 10 stratified folds by target as well as year
folds <- create_folds(pts.comb.half.s$Forest_B, k = 10, seed = s, m_rep = 1, shuffle=FALSE)

# save error scores for each fold
cv_acc_comb <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_acc_comb) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')
# save AUC scores for each fold
cv_auc_comb <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_auc_comb) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')

# object to save variable importance results to
cv_imp <- data.frame(matrix(ncol = length(feats)-1, nrow = 10))
colnames(cv_imp) <- feats[-1]

if (exists("rf.cv.cm.comb")) {
  remove(rf.cv.cm.comb)
}

if (exists("brt.cv.cm.comb")) {
  remove(brt.cv.cm.comb)
}

# loop over each set of train and test folds
for (i in seq_along(folds)) {
  train <- pts.comb.half.s[folds[[i]], ]
  test <- pts.comb.half.s[-folds[[i]], ]
  
  # perform over-sampling
  train <- ovun.sample(Forest_B~., data=train, method='over')$data
  
  # train DT model
  dt.mdl <- decision_tree() %>%
    set_engine("rpart") %>%
    set_mode("classification") %>% 
    fit(Forest_B ~ ., data = train)
  
  # train RF model
  rf.mdl <- randomForest(Forest_B ~ .,
    data = train,
    importance = TRUE,
    ntree = 1000, # n rounds
  )
  
  # train BRT model
  brt.mdl <- boost_tree(
    mode = "classification",
    trees = 1000, #nrounds
  ) %>%
    set_engine("xgboost", verbose=0) %>%
    fit(Forest_B ~ ., data = train)
  
  # Make predictions on the testing data
  dt.p <- as.matrix(dt.mdl %>% predict(pts.nb.b))
  rf.p <- as.matrix(rf.mdl %>% predict(pts.nb.b))
  brt.p <- as.matrix(brt.mdl %>% predict(pts.nb.b))

  # confusion matrices
  dt.cm <- confusionMatrix(table(dt.p, pts.nb.b$Forest_B), positive='1')
  rf.cm <- confusionMatrix(table(rf.p, pts.nb.b$Forest_B), positive='1')
  brt.cm <- confusionMatrix(table(brt.p, pts.nb.b$Forest_B), positive='1')
  
  # add random forest confusion matrix
  if (exists("rf.cv.cm.comb")) {
    rf.cv.cm.comb <- rf.cv.cm.comb + rf.cm$table
  } else {
    rf.cv.cm.comb <- rf.cm$table
  }
  
  # add BRT confusion matrix
  if (exists("brt.cv.cm.comb")) {
    brt.cv.cm.comb <- brt.cv.cm.comb + brt.cm$table
  } else {
    brt.cv.cm.comb <- brt.cm$table
  }
  
  # save accuracy scores from confusion matrices
  cv_acc_comb[i,'Decision Tree'] <- dt.cm$byClass[11]
  cv_acc_comb[i,'Random Forest'] <- rf.cm$byClass[11]
  cv_acc_comb[i,'Boosted Regression Tree'] <- brt.cm$byClass[11]
  
  # save AUC scores from confusion matrices
  pts.nb.b$Forest <- as.numeric(pts.nb.b$Forest_B)
  cv_auc_comb[i,'Decision Tree'] <- auc(as.numeric(dt.p), pts.nb.b$Forest)
  cv_auc_comb[i,'Random Forest'] <- auc(as.numeric(rf.p), pts.nb.b$Forest)
  cv_auc_comb[i,'Boosted Regression Tree'] <- auc(as.numeric(brt.p), pts.nb.b$Forest)
  
  # save feature importance 
  var_imp <- vip::vip(rf.mdl, num_features = length(feats))
  df_imp  <- as.data.frame(var_imp$data)
  df_imp[match(feats, df_imp$Variable),]
  cv_imp[i,] <- as.vector(df_imp$Importance)
}
```

```{r}
s = 123 # setting seed

pts.stack.3.s <- pts.3 %>% dplyr::select(feats)

# 10 stratified folds by target as well as year
folds <- create_folds(pts.stack.3.s$Forest_B, k = 10, seed = s, m_rep = 1, shuffle=FALSE)
# save error scores for each fold
cv_acc_3_nb <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_acc_3_nb) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')
# save AUC scores for each fold
cv_auc_3_nb <- data.frame(matrix(ncol = 3, nrow = 10))
colnames(cv_auc_3_nb) <- c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')

# loop over each set of train and test folds
for (i in seq_along(folds)) {
  train <- pts.stack.3.s[folds[[i]], ]
  test <- pts.stack.3.s[-folds[[i]], ]
  
  # perform over-sampling
  train <- ovun.sample(Forest_B~., data=train, method='over')$data
  
  # train DT model
  dt.mdl <- decision_tree() %>%
    set_engine("rpart") %>%
    set_mode("classification") %>% 
    fit(Forest_B ~ ., data = train)
  
  # train RF model
  rf.mdl <- randomForest(Forest_B ~ .,
    data = train,
    importance = TRUE,
    ntree = 1000, # n rounds
  )
  
  # train BRT model
  brt.mdl <- boost_tree(
    mode = "classification",
    trees = 1000, #nrounds
  ) %>%
    set_engine("xgboost", verbose=0) %>%
    fit(Forest_B ~ ., data = train)
  
  # Make predictions on the testing data
  dt.p <- as.matrix(dt.mdl %>% predict(pts.nb.b))
  rf.p <- as.matrix(rf.mdl %>% predict(pts.nb.b))
  brt.p <- as.matrix(brt.mdl %>% predict(pts.nb.b))

  # confusion matrices
  dt.cm <- confusionMatrix(table(factor(dt.p, levels=0:1), 
      factor(pts.nb.b$Forest_B, levels=0:1)))
  rf.cm <- confusionMatrix(table(factor(rf.p, levels=0:1), 
      factor(pts.nb.b$Forest_B, levels=0:1))) 
  brt.cm <- confusionMatrix(table(factor(brt.p, levels=0:1), 
      factor(pts.nb.b$Forest_B, levels=0:1)))
  
  # save accuracy scores from confusion matrices
  cv_acc_3_nb[i,'Decision Tree'] <- dt.cm$byClass[11]
  cv_acc_3_nb[i,'Random Forest'] <- rf.cm$byClass[11]
  cv_acc_3_nb[i,'Boosted Regression Tree'] <- brt.cm$byClass[11]
  
  # save AUC scores from confusion matrices
  pts.nb.b$Forest <- as.numeric(pts.nb.b$Forest_B)
  cv_auc_3_nb[i,'Decision Tree'] <- auc(as.numeric(dt.p), pts.nb.b$Forest)
  cv_auc_3_nb[i,'Random Forest'] <- auc(as.numeric(rf.p), pts.nb.b$Forest)
  cv_auc_3_nb[i,'Boosted Regression Tree'] <- auc(as.numeric(brt.p), pts.nb.b$Forest)
  
}
```


```{r fig.cap="Figure 8: Comparision of Random Forest model balanced accuracy with and without using combined data from 2020, 2021, and 2023."}
cv_acc_comb$Data <- 'All years'
cv_acc_3_nb$Data <- '2022'

rbind(cv_acc_3_nb, cv_acc_comb) %>%
  gather('Model', 'Accuracy', c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')) %>%
  mutate(Data=factor(Data, levels=c('2022', 'All years'))) %>%
  mutate(Model=factor(Model, levels=c('Decision Tree', 'Random Forest', 'Boosted Regression Tree'))) %>%
  ggplot(aes(x=Model, y=Accuracy, fill=Data)) + geom_boxplot() + ylab('Balanced Accuracy')

cv_auc_comb$Data <- 'All years'
cv_auc_3_nb$Data <- '2022'

rbind(cv_auc_3_nb, cv_auc_comb) %>%
  gather('Model', 'AUC', c('Decision Tree', 'Random Forest', 'Boosted Regression Tree')) %>%
  mutate(Data=factor(Data, levels=c('2022', 'All years'))) %>%
  mutate(Model=factor(Model, levels=c('Decision Tree', 'Random Forest', 'Boosted Regression Tree'))) %>%
  ggplot(aes(x=Model, y=AUC, fill=Data)) + geom_boxplot() + ylab('AUC')
```

```{r}
feats.ordered <- colMedians(cv_imp %>% as.matrix()) %>% sort(decreasing = FALSE) %>% names()

(g3 <- cv_imp %>%
  gather('Feature', 'Value') %>%
  mutate(Value=as.numeric(Value)) %>%
  ggplot(aes(y=factor(Feature, level = feats.ordered), x=Value)) + geom_boxplot() + ylab('Feature') + xlab('Importance'))
```



### Predict Forest Cover For Entire Fire

```{r}
# Function for predicting with BRT model
predict_brt_terra <- function(...){
  library(xgboost)
  p<-predict(...)
  return(as.matrix(as.numeric(p[, 1, drop=T]))) 
}

# for XGB model if needed
xgbpred_terra <- function(model, data) {
  library(xgboost)
  predict(model, newdata=as.matrix(data))
}
```

```{r}
# custom function to perform cross-validated predictions on SpatRaster
predict_cv_terra <- function(f, df, raster, feats, fun=predict) {
  train <- df[f,]
  test <- df[-f,]
  
  # train xgb model
  mdl <- randomForest(Forest_B ~ .,
    data = df,
    importance = TRUE,
    ntree = 1000, # n rounds
  )
  
  # predict on full image
  pred <- terra::predict(raster %>% terra::subset(feats[-1]), model=mdl, fun=predict)
  
  return(pred)
}
```


```{r}
# Train models using both original data plus the non-disturbed data
brt.mdl.comb <- boost_tree(
    mode = "classification",
    trees = 500, #nrounds
  ) %>%
    set_engine("xgboost", verbose=0) %>%
    fit(Forest_B ~ ., data = pts.comb.s)
```


```{r}
##### Riverside ################################################################
rv.stack.0 <- c(
  rast(df.img %>% filter(Year == '2020_pre', Fire == 'Riverside') %>% dplyr::select('Stack') %>% pull()),
  rast(rast(df.img %>% filter(Year == '2020_pre', Fire == 'Riverside') %>% dplyr::select('Stack') %>% pull(), lyrs=1), val=1) %>% categories(value=data.frame(id=1, fire='Riverside'))
  ) %>% terra::subset(c('NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire'))

# Riverside
terra::predict(rv.stack.0, brt.mdl.comb, fun=predict_brt_terra, cores=1, filename=here('forest_cover_normal/Riverside_ForestCover_BRT_2020_Prefire.tif'))
gc()
gc()

remove(rv.stack.0)
gc()
```


```{r}
##### Beachie Creek ################################################################
bc.stack.0 <- c(
  rast(df.img %>% filter(Year == '2020_pre', Fire == 'BeachieCreek') %>% dplyr::select('Stack') %>% pull()),
  rast(rast(df.img %>% filter(Year == '2020_pre', Fire == 'BeachieCreek') %>% dplyr::select('Stack') %>% pull(), lyrs=1), val=2) %>% categories(value=data.frame(id=2, fire='BeachieCreek'))
  ) %>% terra::subset(c('NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire'))

# Beachie Creek
terra::predict(bc.stack.0, brt.mdl.comb, fun=predict_brt_terra, cores=1, filename=here('forest_cover_normal/BeachieCreek_ForestCover_BRT_2020_Prefire.tif'))
gc()
gc()

remove(bc.stack.0)
gc()
```


```{r}
##### Lionshead ################################################################
lh.stack.0 <- c(
  rast(df.img %>% filter(Year == '2020_pre', Fire == 'Lionshead') %>% dplyr::select('Stack') %>% pull()),
  rast(rast(df.img %>% filter(Year == '2020_pre', Fire == 'Lionshead') %>% dplyr::select('Stack') %>% pull(), lyrs=1), val=3) %>% categories(value=data.frame(id=3, fire='Lionshead'))
  ) %>% terra::subset(c('NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire'))

# Lionshead
terra::predict(lh.stack.0, brt.mdl.comb, fun=predict_brt_terra, cores=1, filename=here('forest_cover_normal/Lionshead_ForestCover_BRT_2020_Prefire.tif'))
gc()
gc()

remove(lh.stack.0)
gc()
```


```{r}
##### Holiday Farm ################################################################
hf.stack.0 <- c(
  rast(df.img %>% filter(Year == '2020_pre', Fire == 'HolidayFarm') %>% dplyr::select('Stack') %>% pull()),
  rast(rast(df.img %>% filter(Year == '2020_pre', Fire == 'HolidayFarm') %>% dplyr::select('Stack') %>% pull(), lyrs=1), val=4) %>% categories(value=data.frame(id=4, fire='HolidayFarm'))
  ) %>% terra::subset(c('NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire'))

# Holiday Farm
terra::predict(hf.stack.0, brt.mdl.comb, fun=predict_brt_terra, cores=1, filename=here('forest_cover_normal/HolidayFarm_ForestCover_BRT_2020_Prefire.tif'))
gc()
gc()

remove(hf.stack.0)
gc()
```


```{r}
##### Archie Creek ################################################################
ac.stack.0 <- c(
  rast(df.img %>% filter(Year == '2020_pre', Fire == 'ArchieCreek') %>% dplyr::select('Stack') %>% pull()),
  rast(rast(df.img %>% filter(Year == '2020_pre', Fire == 'ArchieCreek') %>% dplyr::select('Stack') %>% pull(), lyrs=1), val=5) %>% categories(value=data.frame(id=5, fire='ArchieCreek'))
  ) %>% terra::subset(c('NIR', 'NDVI', 'GNDVI', 'BRGHT', 'SDB', 'fire'))

# Archie Creek
terra::predict(ac.stack.0, brt.mdl.comb, fun=predict_brt_terra, cores=1, filename=here('forest_cover_normal/ArchieCreek_ForestCover_BRT_2020_Prefire.tif'))
gc()
gc()

remove(ac.stack.0)
gc()
```


FIN